---
title: "Lab 2-2"
author: "Ollie L. Schilling"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

```{r setup}
knitr::opts_chunk$set(cache = FALSE)
```

## Introduction
This notebook creates an sf object describing SNAP and Medicaid benefit usage in Missouri.

## Dependencies
This notebook requires...

```{r load-packages}
# tidyverse packages
library(dplyr)       # data wrangling

# spatial packages
library(mapview)     # preview geometric data
library(sf)          # spatial tools
library(tidycensus)  # demographic data
library(tigris)      # tiger/line data

# other packages
library(here)       # file path tools
```

## Part 1
### Question 1
This notebook requires both SNAP data and Medicaid data, to be joined.

```{r preview-variables}
acs <- load_variables(2019, "acs5", cache = TRUE) 
```

## Variables to map are:
1. PUBLIC ASSISTANCE INCOME OR FOOD STAMPS/SNAP IN THE PAST 12 MONTHS FOR HOUSEHOLDS
2. MEDICAID/MEANS-TESTED PUBLIC COVERAGE BY SEX BY AGE

### Question 2
This code will download the SNAP necessary data for the specified variables "B19058_001" and "B19058_002". The data will include all counties in Missouri.

```{r download-snap-counties}
snapCounties <- get_acs(geography = "county", year = 2019, state = 29, variables = c("B19058_001", "B19058_002"), output = "wide", geometry = TRUE )
```
```{r preview-snap-counties}
mapview(snapCounties)
```
### Question 3
This code will clean the snapCounties data.
```{r clean-snap}
snapCounties %>%
  rename(total_pop = B19058_001E,
    total_pop_mo = B19058_001M,
    snap = B19058_002E,
    snap_mo = B19058_002M) -> snapCounties
```

## Part 2
### Question 4
This code will download the necessary Medicaid data for the specified variables "C27007_002" and "C27007_012".
```{r download-medicaid}
medicaidCounties <- get_acs(geography = "county", year = 2019, state = 29, variables = c("C27007_002", "C27007_012"), output = "wide", geometry = FALSE)
```
### Question 5
This code will clean the medicaid data.
```{r clean-medicaid}
medicaidCounties %>%
  rename(
    medicaid_male = C27007_002E,
    medicaid_male_mo = C27007_002M,
    medicaid_female = C27007_012E,
    medicaid_female_mo = C27007_012M) %>%
  mutate(medicaid = medicaid_male + medicaid_female) %>% 
  select(-NAME) -> medicaidCounties
```

## Part 3
### Question 6
This code will join data sets together.
```{r join-counties}
services <- left_join(snapCounties, medicaidCounties, by = "GEOID")
```
#Preview the data
```{r preview-counties}
mapview(services, zcol = "medicaid")
```
A description of the results.
