---
title: "Lab-05 Replication Notebook"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

```{r setup}
knitr::opts_chunk$set(cache = FALSE)
```

## Introduction
This is the replication notebook for Lab-05 from the course SOC 4650/5650: Introduction to GISc.

## Load Dependencies
The following code loads the package dependencies for our analysis:

```{r package-load}
# tidyverse packages
library(dplyr)      # data wrangling

# spatial packages
library(mapview)    # preview spatial data
library(sf)         # spatial data tools
library(tidycensus) # data wrangling
library(tigris)     # data wrangling

# other packages
library(here)       # file path tools
```

## Part 1
### Question 1
First, we'll download and preview the variables using the `load_variables()` function from `tidycensus`.

```{r preview-vars}
acs <- load_variables(2019, "acs5", cache = TRUE)
```

The variables we need represent:

  * `"PUBLIC ASSISTANCE INCOME OR FOOD STAMPS/SNAP IN THE PAST 12 MONTHS FOR HOUSEHOLDS"`
  * `"MEDICAID/MEANS-TESTED PUBLIC COVERAGE BY SEX BY AGE"`

### Question 2
First, we'll download the relevant ACS data using `get_acs()`. We get the data for all counties by specifying `"county"` as the geography:

```{r download-snap-counties}
snapCounties <- get_acs(geography = "county", year = 2019, state = 29, 
                        variables = c("B19058_001", "B19058_002"),
                        output = "wide", geometry = TRUE)
```

We can preview our geometric data with `mapview`:

```{r preview-snap-counties}
mapview(snapCounties)
```

### Question 3
Next, we'll clean our data by renaming variables:

```{r clean-snap}
snapCounties %>%
  rename(
    total_pop = B19058_001E,
    total_pop_moe = B19058_001M,
    snap = B19058_002E,
    snap_moe = B19058_002M
  ) -> snapCounties
```

## Part 2
### Question 4
Next, we'll download the relevant ACS data for Medicaid using `get_acs()`:

```{r download-medicaid}
medicaidCounties <- get_acs(geography = "county", year = 2019, state = 29, 
                        variables = c("C27007_002", "C27007_012"),
                        output = "wide", geometry = FALSE)
```

Now we have the number of male and female Medicaid recipients.

### Question 5
Next, we'll tidy up the demographic data, including by renaming variables and summing our male and female Medicaid estimates:

```{r clean-snap-tracts}
medicaidCounties %>%
  rename(
    medicaid_male = C27007_002E,
    medicaid_male_moe = C27007_002M,
    medicaid_female = C27007_012E,
    medicaid_female_moe = C27007_012M
  ) %>%
  mutate(medicaid = medicaid_male + medicaid_female) %>%
  select(-NAME) -> medicaidCounties
```

Now our data are ready to join with our SNAP recipiency data!

## Part 3
### Question 6
Finally, we'll combine our data:

```{r join-counties}
services <- left_join(snapCounties, medicaidCounties, by = "GEOID")
```

To make sure things went correctly, we'll preview our data again:

```{r preview-counties}
mapview(services, zcol = "medicaid")
```

Our data map correctly!
